---
title: "In class Exercise 9"
author: "Dr. Thomas Pollet, Northumbria University (thomas.pollet@northumbria.ac.uk)"
date: '`r format(Sys.Date())` | [disclaimer](http://tvpollet.github.io/disclaimer)'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise... .

1. Calculate a summary table which has the mean/SD of the horse power variable organized by number of gears. (Bonus: export it to .html or Word.)

2. Make a new dataframe called my_cars which contains the columns mpg, hp columns but let the column names be miles_per_gallon and horse_power respectively.

3. Create a new variable in the dataframe called km_per_litre using the mutate function. Note: 1 mpg = 0.425 km/l .

4. Look at the sample_frac() function. Use it to make a new dataframe with a random selection of half the data.

5. Look at the slice function. From the original dataframe select rows 10 to 35.

6. Look at the tibble package and the rownames_to_column function. Make a dataset with just the "Lotus Europa" model. What would be an alternative way of reaching the same goal?

### 1.

```{r}

library(datasets)
cars<-datasets::mtcars
library(dplyr)
grouped<-group_by(cars, gear)
table<-summarise(grouped, Mean=mean(hp), Sd=sd(hp)) %>% mutate_all(funs(round(., 3)))
require(stargazer)
stargazer(table, summary=F, type="html", out= "horsepower.html", header=F)
```

### 2.

Note that I have not reloaded dplyr. Mutate will also get you there but you'd have to then remove the surplus columns.

```{r}
my_cars <- cars %>% select(miles_per_gallon = mpg, horse_power=hp) 
```

### 3.

I have added it to the my_cars dataframe.

```{r}
my_cars <- my_cars %>% mutate(km_per_litre = 0.425*miles_per_gallon)
```

### 4. 

Sliced some rows.

```{r}
my_cars_slice = my_cars %>% slice(10:35)
```

### 5. 

```{r}  
my_cars_sample <- my_cars %>% sample_frac(size = 0.5, replace = FALSE)
```

### 6.

This requires tibble. But if you loaded the tidyverse, it should be in good order.

```{r}
require(tidyverse)
mycars_final = rownames_to_column(mtcars, var = "model")
Lotus_europa <- mycars_final %>% filter(model == "Lotus Europa")
```

Alternatively.

```{r}
Lotus_europa2<- mtcars %>% filter(rownames(mtcars) %in% "Lotus Europa")
```

Or.

```{r}
Lotus_europa3 <- mtcars %>% filter(rownames(mtcars) == "Lotus Europa")
```

You could also make a new variable of row names via mutate.

