---
title: "SEM madness with the Wheaton data"
author: "Dr. Thomas Pollet, Northumbria University (thomas.pollet@northumbria.ac.uk)"
date: '`r format(Sys.Date())` | [disclaimer](http://tvpollet.github.io/disclaimer)'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hypothesised model

# Data

```{r}
require(lavaan)
lower <- '
 11.834,
  6.947,    9.364,
  6.819,    5.091,   12.532,
  4.783,    5.028,    7.495,    9.986,
 -3.839,   -3.889,   -3.841,   -3.625,   9.610,
-21.899,  -18.831,  -21.748,  -18.775,  35.522,  450.288 '

# convert to a full symmetric covariance matrix with names
wheaton.cov <- getCov(lower, names=c("anomia67","powerless67", "anomia71",
                                     "powerless71","education","sei"))

```

## Model

```{r}
require(lavaan)
# the model
wheaton.model = '
  # measurement model
    ses     =~ education + sei
    alien67 =~ anomia67 + powerless67
    alien71 =~ anomia71 + powerless71
 
  # structural model
    alien71 ~ aa*alien67 + ses
    alien67 ~ sa*ses
 
  # correlated residuals
    anomia67 ~~ anomia71
    powerless67 ~~ powerless71

  # Indirect effect
    IndirectEffect := sa*aa
'
```

## Fit the model.


```{r}
require(lavaan)
alienation <- sem(wheaton.model, sample.cov=wheaton.cov, sample.nobs=932)
summary(alienation, fit.measures=T)
```

## Plot

```{r, message=F, warning=F, fig.align="center",fig.height=3.5, fig.width=8}
require(semPlot)
require(qgraph)
semPaths(alienation,style = "lisrel",what="std",  curvePivot = TRUE)
```

# Model without the arrow? 

This model drops the arrow.

```{r}
require(lavaan)
# the model
wheaton.model2 = '
  # measurement model
    ses     =~ education + sei
    alien67 =~ anomia67 + powerless67
    alien71 =~ anomia71 + powerless71
 
  # structural model
    alien71 ~ ses
    alien67 ~ sa*ses
 
  # correlated residuals
    anomia67 ~~ anomia71
    powerless67 ~~ powerless71
'
```

You'll notice that the fit statistics are identical?!

Why is that? We have modelled the covariance now instead of the arrow! In short, we cannot differentiate between double-headed arrow and directed arrow in terms of fit! It is a decision by the researcher to have one or the other but they are equivalent models -- Read more in the sources cited. 

```{r}
require(lavaan)
alienation_without <- sem(wheaton.model2, sample.cov=wheaton.cov, sample.nobs=932)
summary(alienation_without, fit.measures=T)
```

```{r, message=F, warning=F, fig.align="center",fig.height=3.5, fig.width=8}
require(semPlot)
require(qgraph)
semPaths(alienation_without,style = "lisrel",what="std",  curvePivot = TRUE)
```

# Model without covariance between latent constructs

```{r}
require(lavaan)
# the model
wheaton.model3 = '
  # measurement model
    ses     =~ education + sei
    alien67 =~ anomia67 + powerless67
    alien71 =~ anomia71 + powerless71
 
  # structural model
    alien71 ~ ses
    alien67 ~ sa*ses
    
  # Orthogonal (not correlated)  
    alien71 ~~ 0*alien67
 
  # correlated residuals
    anomia67 ~~ anomia71
    powerless67 ~~ powerless71
'
```

```{r}
require(lavaan)
alienation_without2 <- sem(wheaton.model3, sample.cov=wheaton.cov, sample.nobs=932)
summary(alienation_without2, fit.measures=T)
```
```{r, message=F, warning=F, fig.align="center",fig.height=3.5, fig.width=8}
require(semPlot)
require(qgraph)
semPaths(alienation_without2,style = "lisrel",what="std",  curvePivot = TRUE)
```

This model is clearly inferior in terms of fit and also in absolute terms no longer a good fit in RMSEA.

## Model without correlated residuals

```{r}
require(lavaan)
# the model
wheaton.model_no_corr_resid = '
  # measurement model
    ses     =~ education + sei
    alien67 =~ anomia67 + powerless67
    alien71 =~ anomia71 + powerless71
 
  # structural model
    alien71 ~ aa*alien67 + ses
    alien67 ~ sa*ses

  # Indirect effect
    IndirectEffect := sa*aa
'
```

## Fit the model.


```{r}
require(lavaan)
alienation_no_corr_resid <- sem(wheaton.model_no_corr_resid, sample.cov=wheaton.cov, sample.nobs=932)
summary(alienation_no_corr_resid, fit.measures=T)
```

You can tell that this model is substantially worse. But what if we just remove the powerlessness correlated residuals?

```{r}
require(lavaan)
# the model
wheaton.model_only_corr_resid = '
  # measurement model
    ses     =~ education + sei
    alien67 =~ anomia67 + powerless67
    alien71 =~ anomia71 + powerless71
 
  # structural model
    alien71 ~ aa*alien67 + ses
    alien67 ~ sa*ses
    
  # correlated residuals (anomia)
    anomia67 ~~ anomia71

  # Indirect effect
    IndirectEffect := sa*aa
'
```

## Fit the model.

Nearly identical AIC but some improvement in BIC.

```{r}
require(lavaan)
alienation_only_corr_resid <- sem(wheaton.model_only_corr_resid, sample.cov=wheaton.cov, sample.nobs=932)
summary(alienation_only_corr_resid, fit.measures=T)
```

# Session info.

```{r}
sessionInfo()
```

