---
title: "Power simulation for a RI-CLPM"
author: "Prof. Thomas Pollet, Northumbria University, UK"
date: '`r format(Sys.Date())`'
output:
    html_document:
     df_print: paged
     theme: cerulean
     highlight: haddock
     toc: yes
     toc_depth: 3
     toc_float:
      collapsed: false
      smooth_scroll: true
    code_fold: show
bibliography: r-analysis-references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

A paper I reviewed conducted an RI-CLPM, I wanted to verify the proposed sample size for 80% power. The authors proposed _n_ = 400. I thought I would share the code, in case anybody wanted to have a go at a power analysis for RI-CLPM.

# Packages


```{r packages}
library("powRICLPM")
library("tidyverse")
library("future")
library("furrr")
library("progressr")
library("ggplot2")
library("ggthemes")
library("wesanderson")
library("grateful")
library("papaja")
```

```{r}
future::plan(multisession)
```


# Simulation.

## Setup

Values from Manual.

```{r}
Phi <- matrix(c(.4, .1, .15, .3), ncol = 2, byrow = TRUE)
```

```{r}
check_Phi(Phi)
```

Below uses reasonable starting values from the [manual](https://jeroendmulder.github.io/powRICLPM/articles/start.html). 

However, I changed a key path of interest, this is .15 (A on B). (Authors proposed _r_ = .15)

Number of reps is perhaps smallish (but I boosted it from example), you might want to boost it to 10k but for our purpose 1,000 seems sufficient. Note that this takes some time.

I have hidden the warnings error as `lavaan` will print various warnings for individual models.

```{r message=FALSE, warning= F}
out_preliminary <- powRICLPM(target_power = 0.8,search_lower = 200, search_upper = 1000, search_step = 50, time_points = c(3, 4, 5), ICC = c(0.3, 0.5, 0.7), RI_cor = 0.3, Phi = Phi, within_cor = 0.3, reps = 1000, seed = 1234567)
```

## Output

```{r}
summary(out_preliminary)
```

A on B --> .15 

The graph could be further beautified, but I have left it here. And perhaps Tufte's theme is not ideal.

```{r}
plot(out_preliminary, parameter = "wB2~wA1") + theme_tufte() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  +
  scale_color_manual(values = wes_palette("GrandBudapest1")) + scale_fill_manual(values = wes_palette("GrandBudapest1")) + guides(shape = "none")
```

This is a weaker effect.

B on A --> .1

```{r}
plot(out_preliminary, parameter = "wA2~wB1") + theme_tufte() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  +
  scale_color_manual(values = wes_palette("GrandBudapest1")) + scale_fill_manual(values = wes_palette("GrandBudapest1"))  + guides(shape = "none")
```

# Conclusion

The above simulations suggest that _n_ = 400 does not seem to satisfy 80% power, under any scenario. The above code can be modified to other scenarios. Note that as described by the package author, there are assumptions made (e.g., Multivariate normality, etc) -- see the paper by [Mulder (2022)](https://doi.org/10.1080/10705511.2022.2122467). So, a broader range of scenarios might need to be explored for a particular design.

# Package citations

```{r}
r_refs(file = "r-analysis-references.bib")
my_citation <- cite_r(file = "r-analysis-references.bib")
```

`r my_citation`

# SessionInfo

```{r}
sessionInfo()
```
